Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-08 paper/paper.tex
fix: semantic hypothesis 1.0

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extractable from the environment
\citep{bar2009proactive}.
In this way, the encoding of complex environments could be facilitated
by relative similarity associations between states.
%
Going much beyond human language ifself,
semantic building blocks provide the basis to
mentally envision non-existent scenarios
that would improve optimal behavior in the environment
by simulating future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and are associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account explains the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, social knowledge, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
as these necessitate abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
These scene generation processes could contribute to interpreting the
present and foretelling the future.
This view is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

\begin{itemize}
  \item Both the semantic hypothesis and MDPs expose mechanisms of
  how alternative decision trees could be mentally explored.
  \item In both semantic hypothesis and MDPs,
  there is no evidence to indicate that predictions of various
  levels of complexity, abstraction, timescale and purpose
  use mechanisms that are qualitatively different. This concurs with
  DMN activity increases across time, space, and content domains
  demonstrated in many neuroimaging studies. The semantic hypothesis
  and RL account provide explanations why hippocampus lesion does
  not only impair retrieving memories, but also hypothetical and future
  thinking \citep{hassabis2007patients}.
  \item The notion of semantic or knowledge association is
  incorporated into the MDP as the Markov property,
  that is, the current state directly results from the
  agent's history of states and actions. The learned
  value matrix and action transition schedules drive
  stimulus processing and action choice in the present.
  \item Mental scene construction has been proposed by some authors
  \citep{buckner2007self} to imply a distinction between
    environmental perception and internally generated mind-wandering\footnote{
      \textbf{XXX Elvis: This is more than just a footnote. Make this part of the paper
      and factorize the lot. Also you probably need a citation here!!!}
  ``A computational model of how such a process might be structured
  is far from being defined, but it will probably require a form of
  regulation by which perception of the current world is suppressed
  while simulation of possible alternatives are constructed,
  followed by a return to perception of the present.''\citep{buckner2007self}}.
  The MDPs naturally integrate the former egocentric
  (more related to current action, state and reward) and the later
  allocentric (more related to past and future actions, states, and rewards)
  angles on the world in a same optimization problem.
  \item The semantic account of DMN function does not offer
  a mechanistic explanation how explicit world knowledge and semantic analogies thereof
  lead to prediction of future actions and states.
  \item In contrast to existing accounts on semantics and
  mental scene construction, the random and creative aspects of DMN function
  are explained in MDPs by the advantages of stochastic optimization.
  \item The semantic hypothesis does not explain why memory recall
  for scene construction in humans is typically fragmentary and noisy
  instead of accurate and reliable.
  Yet, the MDP framework provides an algorithmic explanation in that
  stochasticity of the parameter space explored
  by the Monte Carlo solvers achieves better fine-tuning of the
  action policies and estimation of expected reward outcomes.
  That is, the purposeful stochasticity of policy and value estimation
  in MDPs provides a candidate explanation for why humans
  have evolved imperfect noisy memories
  as the more advantageous adaptation.
  \item While both semantic and MDP account propose memory-based internally
  generated information for probabilistic mental models of action outcomes,
  only MDPs render explicit the grounds on which the final action is
  eventually chosen (i.e., the estimated cumulative reward).
  \item In sum, episodic scene construction according to the semantic
  hypothesis is lacking an explicit time and incentive model;
  neither does it explain how randomness of human mental experience
  can be beneficial.
\end{itemize}

>>>>

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extracted from the environment
\citep{bar2009proactive}.
The encoding of complex environments could thus be facilitated
by relative similarity associations between states.
%
Going beyond human language itself,
semantic building blocks may provide the basis for
mentally envisioning non-existent scenarios
that would optimize behavior in the environment
by constantly simulating possible future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account can explain the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, semantic associations, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation.
These neural processes are likely to
build on abstract world knowledge and abstract assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
Such scene generation processes could contribute to interpreting the
present and foretelling the future.
Mental scene construction has been proposed
to imply a distinction between
environmental perception and internally generated mind-wandering
\citep{buckner2007self}.
These authors stated that
``A computational model [...] will probably require a form of
regulation by which perception of the current world is suppressed
while simulation of possible alternatives are constructed,
followed by a return to perception of the present.''.
The semantic hypothesis is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

In comparison,
both the semantic hypothesis and MDPs expose mechanisms of
how action considerations could be mentally explored.
In both accounts,
there is also no evidence to indicate that predictions of various
levels of complexity, abstraction, timescale and purpose
use mechanisms that are qualitatively different. This concurs with
DMN activity increases across time, space, and content domains
demonstrated in many neuroimaging studies \citep{spreng2009common}.
The semantic hypothesis
and MDP account provide explanations why hippocampus lesion does
not only impair retrieving memories, but also hypothetical and future
thinking \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-based internally
generated information for probabilistic mental models of action outcomes,
MDPs render explicit the grounds on which the final action is
eventually chosen, namely, the estimated cumulative reward.
\textbf{The notion of semantic knowledge associations is
incorporated into the MDP as the Markov property,
that is, the current state directly results from the
agent's history of states and actions. The learned
value matrix and action transition schedules drive
stimulus processing and action choice in the present.}
The MDPs naturally integrate the former egocentric
(more related to current action, state and reward) and the later
allocentric (more related to past and future actions, states, and rewards)
angles on the world in a same optimization problem.
Further,
the semantic account of DMN function does not offer
a mechanistic explanation how explicit world knowledge and semantic analogies thereof
lead to prediction of future actions and states.
The semantic hypothesis does not explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Yet, the MDP framework provides an algorithmic explanation in that
stochasticity of the parameter space explored
by the Monte Carlo solvers achieves better fine-tuning of the
action policies and estimation of expected reward outcomes.
That is, the purposeful stochasticity of policy and value estimation
in MDPs provides a candidate explanation for why humans
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, episodic scene construction according to the semantic
hypothesis is lacking an explicit time and incentive model;
neither does it explain how randomness of human mental experience
can be beneficial.
